Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-14 paper/sec-analysis.tex
implementing Daniela's comments

  These subjets are constructed by reclustering the constituents
  of a large-$R$ jet, again with  anti-$k_T$ algorithm,
  but this time with a smaller radius parameter, namely
  $R=0.3$~\cite{Aad:2015uka}.
 %
  These small-$R$ subjets will be the main input for
  $b$-tagging in the boosted category.
  %
  They are required to satisfy $p_T > 50$~GeV and $|\eta|<2.5$.
  %
\end{itemize}

In addition, we have
also explored the possible improvements
in the analysis from the use
of variable-$R$ jets~\cite{Krohn:2009zg}, finding
a similar performance as in the case of fixed-$R$ jets.
%
This is explained by the fact that
variable-$R$ jets are more suitable when
the degree of boost of the final state being reconstructed can span
a wide range~\cite{Cacciari:2008gd}, such as in heavy resonance
searches~\cite{Aad:2015fna},
where where the degree of boost of the final state
is {\it a priori} unknown.

For the   boosted and intermediate categories,
which involve the use of large-$R$ jets,
we will use jet substructure variables~\cite{Salam:2009jx,Aad:2013gja} to
improve the significance of the discrimination between signal and background
events in the MVA.
%
In particular we will consider the following jet
substructure variables:
\begin{itemize}
\item The $k_T$-splitting scale~\cite{Butterworth:2002tt,Butterworth:2008iy}.

@@

  The $N$-subjettiness variables $\tau_N$ are defined by clustering the constituents
  of a jet with the exclusive $k_t$ algorithm~\cite{Catani:1993hr}
  and requiring that $N$ subjets are found,
  \be
  \tau_N \equiv \frac{1}{d_0} \sum_k p_{T,k}\cdot {\rm min}\lp \delta R_{1k}, \ldots,
  \delta R_{Nk}\rp \, , \qquad d_0\equiv \sum_k p_{T,k}\cdot R \, ,
  \ee
  with $p_{T,k}$ is the $p_T$ of the constituent particle $k$ and $\delta R_{ik}$ the distance from
  subjet $i$ to constituent $k$.
  %
  In this work we use as input to the MVA the ratio of 2-subjettiness to 1-subjettiness
  \be
  \label{eq:tau21}
\tau_{21} \equiv \frac{\tau_2}{\tau_1} \, ,
  \ee
  which provides good discrimination power
  between QCD jets and jets arising from the decay of
  a heavy resonance.

@@

    Large-$R$ jets are $b$-tagged by
    ghost-associating anti-$k_T$ $R=0.3$ (AKT03)
    subjets to the original large-$R$
    jets~\cite{Cacciari:2007fd,Aad:2013gja,
      ATLAS-CONF-2014-004,Aad:2015uka}.
    %
    A large-$R$ jet will considered $b$-tagged if both
    the leading and subleading AK03 subjets, where the ordering
    is done in the subjet $p_T$, are both individually $b$-tagged,
    with the same criteria as the small-$R$ jets.
    %
     Therefore, a large-$R$ jet where the two leading
    subjets have at least one $b$-quark will be tagged
    with probability $f_b^2$.

@@

The basic idea of the present analysis is similar to the
scale-invariant resonance tagging introduced
in Ref.~\cite{Gouzevitch:2013qca}.
%
Rather than restricting to a specific event topology,
we aim to consistently combine the information from
the three possible topologies: boosted, intermediate and
resolved.
%
The optimal cuts for each category will be determined
separately, and then
a exclusive categorisation is defined starting from the category
with the highest signal significance.
%
An advantage of our approach
the robustness under variations of
the underlying production model of Higgs pairs is modified,
for instance in the case of
BSM dynamics, which can substantially increase
the degree of boost of the final state.

These three categories are defined as follows:
\begin{itemize}
\item {\it Boosted category}.

@@

Once a Higgs boson candidate has been identified,
its invariant mass is required to lie within a fixed window
around the nominal Higgs boson mass of $m_h= 125$
GeV.
%
In particular we require the condition
\be
\label{higgsmasswindow}
|m_{h,j} - 125| < 40~{\rm GeV} \, ,\, j=1,2 \, ,
\ee
where $m_{h,j}$ is the invariant mass of each of the two reconstructed  Higgs candidates.
%
This cut is substantially looser as the corresponding
cut used in the typical ATLAS and CMS $h\to b\bar{b}$
analyses~\cite{Aad:2012gxa,Chatrchyan:2013zna}: the motivation
for this choice is that the MVA will take care of the optimisation of this,
as well as of other cuts that are not explicitly performed.
%
Only events where the two Higgs candidates satisfy
Eq.~(\ref{higgsmasswindow}) are classified as signal events.
%

These three categories are not exclusive:
a given event can be assigned to more than one category, for
example, satisfying the requirements of both the intermediate
and resolved
categories at the same time.
%
The exception is the boosted and intermediate categories, which have
orthogonal jet selection requirements.
%
The separate optimisation of the three categories needs to be performed in
a inclusive analysis, and then we introduce a
 priorisation starting with the category of highest signal significance.

 This has been achieved as follows.
 %
First of all we perform an inclusive analysis, and optimise the
signal significance
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that the category with higher significance turns out to be
the boosted one,
followed by the intermediate and the resolved ones, the latter two
with similar significance.
%
Therefore, for the final analysis, we turn to an exclusive approach:
if an event satisfies the boosted requirements, it is assigned to
this category, else we check if it suits the intermediate
requirements, and failing also this, the event is attempted to
be classified in the
resolved category.
%
The resulting exclusive event samples are then separately processed
through the MVA, allowing a consistent combination
of the significance of the three event categories.

@@

We now motivate the values of
kinematic cuts applied to the different categories, 
comparing representative kinematic distributions between
signal and background events.
%
First to all, we present results without PU, and in the next
subsection we discuss the impact of PU
on the description of the various kinematic
distributions.
%
In the following, all
distributions will be normalized to their total integral.

%
To begin with,
in Fig.~\ref{fig:cutplots1} we show
the $p_T$ distributions
of the
  leading and subleading large-$R$ jets in the boosted category, comparing
  the shapes of the distribution in signal and background events.
  %
  We see that the background distribution
  falls off rather steeply than the signal at large $p_T$
  for the two Higgs candidates.
  %
  On the other hand, the cut in $p_T$ cannot be too strong to avoid
  a substantial degradation of signal selection efficiency,
  specially taking into account the subleading large-$R$ jet.
  %
  This comparison justifies the cut of $p_T \ge 200$ GeV
  for the large-$R$ jets that we impose in the boosted category.

@@

Another selection requirement for the boosted category is that the two
leading AKT03 subjets of the large-$R$ jet must be relatively hard,
in particular they should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig.~\ref{fig:cutplots22}
we show the distribution in $p_T$ of the leading
and subleading AKT03 subjets in the subleading large-$R$ jet in events
corresponding to the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background.
%
On the other hand, considering the subleading AKT03 subjet,
this cut in its  $p_T$
cannot be too harsh, to maintain a high signal selection
efficiency.
%
Therefore,
as for the previous distribution, also here 
a compromise between suppressing backgrounds but keeping a large fraction of
signal events is crucial.

@@

Turning to the resolved category, an important aspect to account
in the corresponding selection
cuts is the fact that the $p_T$ distribution
of the four leading small-$R$ jets of the event can be relatively soft,
specially for the subleading jets.
%
As noted in~\cite{deLima:2014dta}, this is a consequence of the fact
that in general the boost from the Higgs decays is moderate.
%
These considerations lead to the requirement that the $p_T$ cut
for the small-$R$ jets cannot be too strong.
%
In Fig.~\ref{fig:cutplots23}
we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events: we observe that both
distributions are peaked at $p_T \le 50$ GeV, with the signal distribution
eventually becoming dominant at large $p_T$.
%
The feasibility to being able to trigger on four small-$R$ jets with a relatively
soft $p_T$ distribution is one of the experimental challenges of
exploiting the resolved category in this final state,
and hence the requirement that $p_T \ge 40$ GeV for
the small-$R$ jets.
%
In  Fig.~\ref{fig:cutplots23} we also show the
rapidity distribution of the the small-$R$
jets in the resolved category: as well know, the production
is mostly central (more in the case
of signal events, since backgrounds are dominated by
QCD $t$-channel exchange) and thus the standard rapidity cuts
are relatively harmless.

@@

One of the selection cuts with highest discrimination power is the requirement
that the invariant mass of the Higgs candidate (di)jets must lie within a window
around the nominal Higgs value, Eq.~(\ref{higgsmasswindow}).
%
In Fig.~\ref{fig:mHHinv} we show the invariant mass
of the leading reconstructed Higgs candidates, before the Higgs mass window
cut 
  is applied, for the resolved and boosted categories.
%
While the signal distribution is of course peaked at the
nominal Higgs mass, the background distributions
show no particular
structure.
%
The
width of the Higgs mass peak is driven both from QCD effects,
such as initial-state radiation (ISR)
and out-of-cone radiation, as well
as from the four-momenta smearing applied to final state particles.
%

@@

Another important kinematic distribution of this process is the invariant mass
of the di-Higgs system.
%
This is a direct measure of the boost of the system,
which in  BSM scenarios can be substantially
enhanced, for instance due to
of specific $d=6$ EFT operators~\cite{Azatov:2015oxa}.
%
One important advantage of the $b\bar{b}b\bar{b}$ final state for
di-Higgs production is that it significantly increases the reach
in $m_{hh}$ as compared to other channels with smaller branching
ratios,
such as $2b2\gamma$ or $2b2\tau$.
%
In Fig.~\ref{fig:mhh} we show the invariant mass distribution of the
reconstructed Higgs pairs,
comparing the resolved and the boosted categories.

@@

In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as compared
to the background,
and thus one expects that cutting in $m_{hh}$ would help signal
discrimination: we will verify this using the MVA.
%
For the boosted category the overall trend of the $m_{hh}$ distribution
is different because of the jet selection cuts, with the
distribution now peaking at higher values of the invariant mass.
%
In this case, signal and background distributions
look reasonably similar.
%
Note that at parton-level the $m_{hh}$ distribution
for signal events has a kinematic
cut-off at $m_{hh}^{\rm min}=250$ GeV, which is smeared due
to 
the parton shower and to detector resolution effects.
%

In Fig.~\ref{fig:pthh} we show the transverse momentum of
the di-Higgs system, $p_T^{hh}$,
for the resolved and boosted categories.
%
Again we see that the background has a steeper $p_T^{hh}$ distribution
that the signal, in both categories, thus this variable
should provide additional discrimination power, and therefore
this distribution will be another of the inputs for the MVA.
%
In our LO simulation the $p_T^{hh}$ distribution is generated
by the parton shower, and and improved theoretical
description would require
either the matching with higher-multiplicity
matrix elements~\cite{Maierhofer:2013sha} or
the full NLO calculation~\cite{Frederix:2014hta},
%
Nevertheless, as we will show,
the MVA shows only limited sensitivity to this variable, so its
modelling is not crucial in this specific application.

@@

We now study the discrimination power to separate signal
and background events contained in
jet substructure
variables.
%
In Fig.~\ref{fig:mva_substructure_1}
we show the distributions of representative
 substructure variables for the boosted category: the
$k_t$ splitting scale $\sqrt{d_{12}}$, Eq.~(\ref{eq:ktsplitting}), 
the ECF ratio $C_2^{(\beta)}$,
Eq.~(\ref{eq:c2}), and
the 2--to--1 subjettiness ratio $\tau_{12}$, Eq.~(\ref{eq:tau21}),
all for the leading
Higgs candidates, as we as also $\tau_{12}$ for the subleading
Higgs candidates.

@@

From Fig.~\ref{fig:mva_substructure_1}
we can observe how for these substructure variables the shapes of the signal
and background distributions reflect
the inherent differences in the internal structure of jets
between QCD jets and jets originated from the Higgs decays.
%
Crucially, signal and background distributions peak
in rather
different regions: for example, the $k_t$ splitting scale $\sqrt{d_{12}}$
peaks around 80 GeV (40 GeV) for signal (background) events, while
the distribution of the
ECF ratio $C_2^{(\beta)}$ is concentrated at small values
for signal but is much broader for background events.
%
From Fig.~\ref{fig:mva_substructure_1} we also see
the distributions of the subjettiness ratio $\tau_{12}$ are
reasonably similar
for both the leading and in the subleading jets, therefore the information
contained in
this (and other substructure variables)
from each of the two Higgs candidates  will be to some
extent redundant.

Therefore, while we do not explicitly have cuts in the substructure
variables, their discrimination power between signal and
background events is the motivation to include them
as input to the MVA, as we will explain in Sect.~\ref{sec:mva}.

>>>>

  These subjets are constructed by reclustering the constituents
  of a large-$R$ jet, again with the
  anti-$k_T$ algorithm,
  but this time with a smaller radius parameter, namely
  $R=0.3$~\cite{Aad:2015uka}.
 %
  These small-$R$ subjets will be the main input for
  $b$-tagging in the boosted category.
  %
  They are required to satisfy $p_T > 50$~GeV and $|\eta|<2.5$.
  %
\end{itemize}

In addition, we have
also explored the possible improvements
in the analysis from the use
of variable-$R$ jets~\cite{Krohn:2009zg}, finding
a similar performance as in the case of fixed-$R$ jets.
%
This is explained by the fact that
variable-$R$ jets are more suitable when
the degree of boost of the final state being reconstructed can
differ substantially~\cite{Cacciari:2008gd},
such as in heavy resonance
searches~\cite{Aad:2015fna},
where the degree of boost of the final state
is {\it a priori} unknown.

For the   boosted and intermediate categories,
which involve the use of large-$R$ jets,
we use jet substructure variables~\cite{Salam:2009jx,Aad:2013gja} to
improve the significance of the discrimination between signal and background
events in the MVA.
%
In particular we  consider the following jet
substructure variables:
\begin{itemize}
\item The $k_T$-splitting scale~\cite{Butterworth:2002tt,Butterworth:2008iy}.

@@

  The $N$-subjettiness variables $\tau_N$ are defined by clustering the constituents
  of a jet with the exclusive $k_t$ algorithm~\cite{Catani:1993hr}
  and requiring that $N$ subjets are found,
  \be
  \tau_N \equiv \frac{1}{d_0} \sum_k p_{T,k}\cdot {\rm min}\lp \delta R_{1k}, \ldots,
  \delta R_{Nk}\rp \, , \qquad d_0\equiv \sum_k p_{T,k}\cdot R \, ,
  \ee
  with $p_{T,k}$ is the $p_T$ of the constituent particle $k$ and $\delta R_{ik}$ the distance from
  subjet $i$ to constituent $k$.
  %
  In this work we use as input to the MVA the ratio of 2-subjettiness to 1-subjettiness
  \be
  \label{eq:tau21}
\tau_{21} \equiv \frac{\tau_2}{\tau_1} \, ,
  \ee
  which provides good discrimination 
  between QCD jets and jets arising from the decay of
  a heavy resonance.

@@

    Large-$R$ jets are $b$-tagged by
    ghost-associating anti-$k_T$ $R=0.3$ (AKT03)
    subjets to the original large-$R$
    jets~\cite{Cacciari:2007fd,Aad:2013gja,
      ATLAS-CONF-2014-004,Aad:2015uka}.
    %
    A large-$R$ jet is considered $b$-tagged if both
    the leading and subleading AK03 subjets, where the ordering
    is done in the subjet $p_T$, are both individually $b$-tagged,
    with the same criteria as the small-$R$ jets.
    %
     Therefore, a large-$R$ jet where the two leading
    subjets have at least one $b$-quark will be tagged
    with probability $f_b^2$.

@@

The present analysis follows a strategy similar to the
scale-invariant resonance tagging of Ref.~\cite{Gouzevitch:2013qca}.
%
Rather than restricting to a specific event topology,
we aim to consistently combine the information from
the three possible topologies: boosted, intermediate and
resolved, with the optimal cuts for each category being determined
separately.
%
This approach is robust
under variations of
the underlying production model of Higgs pairs is modified,
for instance in the case of
BSM dynamics, which can substantially increase
the degree of boost of the final state.

The three categories are defined as follows:
\begin{itemize}
\item {\it Boosted category}.

@@

Once a Higgs boson candidate has been identified,
its invariant mass is required to lie within a fixed window
around the nominal Higgs boson mass of $m_h= 125$
GeV.
%
In particular we require the condition
\be
\label{higgsmasswindow}
|m_{h,j} - 125| < 40~{\rm GeV} \, ,\, j=1,2 \, ,
\ee
where $m_{h,j}$ is the invariant mass of each of the two reconstructed  Higgs candidates.
%
This cut is substantially looser than the corresponding
cut used in the typical ATLAS and CMS $h\to b\bar{b}$
analyses~\cite{Aad:2012gxa,Chatrchyan:2013zna}.
%
The motivation
for this choice 
is that further improvements of the
signal significance will be obtained using an MVA.
%
Only events where the two Higgs candidates satisfy
Eq.~(\ref{higgsmasswindow}) are classified as signal events.
%

These three categories are not exclusive:
a given event can be assigned to more than one category, for
example, satisfying the requirements of both the intermediate
and resolved
categories at the same time.
%
The exception is the boosted and intermediate categories, which have
orthogonal jet selection requirements.
%
The optimization of the analysis
of the three categories is performed inclusively, and
then events are exclusively assigned to specific category
starting with the one with the highest signal significance.

The separate optimisation of the three categories needs to be performed in
a inclusive analysis, and then we introduce a
 priorisation starting with the category of highest signal significance.

 This has been achieved as follows.
 %
First of all we perform an inclusive analysis, and optimise the
signal significance
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that the category with higher significance is
the boosted one,
followed by the intermediate and the resolved ones, the latter two
with similar significance.
%
Therefore, for the final analysis, we turn to an exclusive approach:
if an event satisfies the boosted requirements, it is assigned to
this category, else we check if it suits the intermediate
requirements.
%
It the event also fails the intermediate category
requirements, we
then check if it pass the resolved selection criteria.
%
The resulting exclusive event samples are then separately processed
through the MVA, allowing a consistent combination
of the significance of the three event categories.

@@

We now motivate the values of
kinematic cuts applied to the different categories, 
comparing representative kinematic distributions between
signal and background events.
%
We first present results without PU, and then
discuss
the impact of PU
on the description of the various kinematic
distributions.
%
In the following, all
distributions will be normalized to their total integral.

In Fig.~\ref{fig:cutplots1} we show
the $p_T$ distributions
of the
  leading and subleading large-$R$ jets in the boosted category, comparing
  the shapes of the distribution in signal and background events.
  %
  We observe that the background distribution
falls off more rapidly as a function of $p_T$ than the di-Higgs signal.
  %
  On the other hand, the cut in $p_T$ cannot be too strong to avoid
  a substantial degradation of signal selection efficiency,
  specially taking into account the subleading large-$R$ jet.
  %
  This comparison justifies the cut of $p_T \ge 200$ GeV
  for the large-$R$ jets that we impose in the boosted category.

@@

Another selection requirement for the boosted category is that the two
leading AKT03 subjets of the large-$R$ jet must be relatively hard,
in particular they should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig.~\ref{fig:cutplots22}
we show the distribution in $p_T$ of the leading
and subleading AKT03 subjets in the subleading large-$R$ jet in events
corresponding to the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background.
%
On the other hand, considering the subleading AKT03 subjet,
this cut in $p_T$
cannot be too harsh, to maintain a high signal selection
efficiency.
%
Therefore,
as for the previous distribution, also here 
a compromise between suppressing backgrounds but keeping a large fraction of
signal events is crucial.

@@

Turning to the resolved category, an important aspect to account
in the corresponding selection
cuts is the fact that the $p_T$ distribution
of the four leading small-$R$ jets of the event can be relatively soft,
specially for the subleading jets.
%
As noted in~\cite{deLima:2014dta}, this is a consequence of the fact
that in general the boost from the Higgs decays is moderate.
%
These considerations lead to the requirement that the $p_T$
selection
for the small-$R$ jets cannot be too large.
%
In Fig.~\ref{fig:cutplots23}
we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events: we observe that both
distributions are peaked at $p_T \le 50$ GeV, with the signal distribution
eventually becoming dominant at large $p_T$.
%
The feasibility to being able to trigger on four small-$R$ jets with a relatively
soft $p_T$ distribution is one of the experimental challenges for
exploiting the resolved category in this final state,
and hence the requirement that $p_T \ge 40$ GeV for
the small-$R$ jets.
%
In  Fig.~\ref{fig:cutplots23} we also show the
rapidity distribution of the the small-$R$
jets in the resolved category.
%
As expected, the production
is mostly central, more in the case
of signal events, since backgrounds are dominated by
QCD $t$-channel exchange, and thus the
selection criteria on the jet rapidity  are highly efficient.

@@

One of the selection cuts with highest discrimination power is the requirement
that the invariant mass of the Higgs candidate (di)jets must lie within a window
around the nominal Higgs value, Eq.~(\ref{higgsmasswindow}).
%
In Fig.~\ref{fig:mHHinv} we show the invariant mass
of the leading reconstructed Higgs candidates, before the Higgs mass window
selection
  is applied, for the resolved and boosted categories.
%
While the signal distribution is of course peaked at the
nominal Higgs mass, the background distributions
show no particular
structure.
%
The
width of the Higgs mass peak is driven both from QCD effects,
such as initial-state radiation (ISR)
and out-of-cone radiation, as well
as from the four-momenta smearing applied to final state particles.
%

@@

Another important kinematic distribution for
this process is the invariant mass
of the di-Higgs system.
%
This is a direct measure of the boost of the system,
which in  BSM scenarios can be substantially
enhanced, for instance due to
specific $d=6$ EFT operators~\cite{Azatov:2015oxa}.
%
One important advantage of the $b\bar{b}b\bar{b}$ final state for
di-Higgs production is that it significantly increases the reach
in $m_{hh}$ as compared to other channels with smaller branching
ratios,
such as $2b2\gamma$ or $2b2\tau$.
%
In Fig.~\ref{fig:mhh} we show the invariant mass distribution of the
reconstructed Higgs pairs,
comparing the resolved and the boosted categories.

@@

In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as compared
to the background,
and thus one expects that cutting in $m_{hh}$ would help signal
discrimination: we will verify this using the MVA.
%
For the boosted category the overall trend of the $m_{hh}$ distribution
is different because of the jet selection criteria, and the
distribution now peaks at higher values of the invariant mass.
%
In this case, signal and background distributions
look similar.
%
Note that at parton-level the $m_{hh}$ distribution
for signal events has a kinematic
cut-off at $m_{hh}^{\rm min}=250$ GeV, which is smeared due
to 
the parton shower and to detector resolution effects.
%

In Fig.~\ref{fig:pthh} we show the transverse momentum of
the di-Higgs system, $p_T^{hh}$,
for the resolved and boosted categories.
%
Again we see that the background has a steeper $p_T^{hh}$ distribution
that the signal, in both categories, thus this variable
should provide additional discrimination power, and therefore
this distribution is one of the inputs for the MVA.
%
In our LO simulation the $p_T^{hh}$ distribution is generated
by the parton shower, and improved theoretical
description would require
either the matching with higher-multiplicity
matrix elements~\cite{Maierhofer:2013sha} or
the full NLO calculation~\cite{Frederix:2014hta},
%
Nevertheless, as we will show,
the MVA shows only limited sensitivity to this variable, so its
modelling is not crucial in this specific application.

@@

We now study the discrimination power to separate signal
and background events contained in
jet substructure
variables.
%
In Fig.~\ref{fig:mva_substructure_1}
we show the distributions of representative
 substructure variables for the boosted category: the
$k_t$ splitting scale $\sqrt{d_{12}}$, Eq.~(\ref{eq:ktsplitting}), 
the ECF ratio $C_2^{(\beta)}$,
Eq.~(\ref{eq:c2}), and
the 2--to--1 subjettiness ratio $\tau_{12}$, Eq.~(\ref{eq:tau21}),
all for the leading
Higgs candidates, and also $\tau_{12}$ for the subleading
Higgs candidates.

@@

From Fig.~\ref{fig:mva_substructure_1}
we can observe how for these substructure variables the shapes of the signal
and background distributions reflect
the inherent differences in the internal structure of jets
between QCD jets and jets originated from the Higgs decays.
%
Signal and background distributions peak
in rather
different regions: for example, the $k_t$ splitting scale $\sqrt{d_{12}}$
peaks around 80 GeV (40 GeV) for signal (background) events, while
the distribution of the
ECF ratio $C_2^{(\beta)}$ is concentrated at small values
for signal but is much broader for background events.
%
From Fig.~\ref{fig:mva_substructure_1} we also see
the distributions of the subjettiness ratio $\tau_{12}$ are
reasonably similar
for both the leading and the subleading jets.
%
Therefore, while we do not explicitly have cuts in the substructure
variables, their discrimination power between signal and
background events is the motivation to include them
as input to the MVA, as discussed in Sect.~\ref{sec:mva}.

### 2015-12-14 paper/sec-intro.tex
implementing Daniela's comments

Analogously to single Higgs production~\cite{Dittmaier:2012vm}, 
in the SM the dominant mechanism for the production of a pair of
Higgs bosons at the LHC is 
gluon fusion (see Refs.~\cite{baglio,Frederix:2014hta} and
references therein).
%
For a center-of-mass energy $\sqrt{s} = 14\,$TeV, the
next-to-next-to-leading order (NNLO)
total cross section is approximately $40\,$fb~\cite{deFlorian:2013jea},
which is increased by a further few percent once
next-to-next-to-leading logarithmic
(NNLL) corrections
are accounted for~\cite{deFlorian:2015moa}.
%
Feasibility studies in the case of a SM-like Higgs boson
in gluon fusion at the LHC have been performed for different final states, including
$b\bar b\gamma\gamma$~\cite{Baur:2003gp,Barger:2013jfa},
$b\bar{b}\tau^+\tau^-$~\cite{Baur:2003gpa,Barr:2013tda,Dolan:2012rv,Dolan:2013rja},
$b\bar{b}W^+W^-$~\cite{Dolan:2012rv,Papaefstathiou:2012qe} and
$b\bar{b}b\bar{b}$~\cite{Baur:2003gpa,Dolan:2012rv,Wardrope:2014kya,deLima:2014dta,Barger:2013jfa}.
%
While these studies differ in their quantitative conclusions,
a consistent picture emerges 
that the ultimate precision in the determination of the Higgs trilinear
coupling $\lambda$ requires the full integrated luminosity
of the HL-LHC, $\mathcal{L}=3$ ab$^{-1}$,
and will rely on the combination of all relevant final states.

@@

While the SM production rates for Higgs
pairs are small, these can be substantially
enhanced in a variety of BSM scenarios.
%
Feasibility studies of Higgs pair production in New Physics
models have been performed in a number of different frameworks,
including Effective Field
Theories (EFTs) with higher-dimensional
operators and anomalous 
Higgs couplings~\cite{Nishiwaki:2013cma,Dall'Osso:2015aia,Azatov:2015oxa,Liu:2014rba,Goertz:2014qta,He:2015spf,Grober:2015cwa}, resonant production
in models such as extra dimensions~\cite{Gouzevitch:2013qca,Cooper:2013kia,No:2013wsa,Wen-Juan:2015gqg}, or Supersymmetry and
Two Higgs Doublet models (2HDMs)~\cite{Belyaev:1999kk,Han:2013sga,Hespel:2014sla,Wu:2015nba,Cao:2014kya,Ellwanger:2013ova}.
%
Since BSM dynamics modify
the kinematic distributions of the Higgs decay products, for
instance boosting the di-Higgs system,
different analysis strategies  might be required for BSM
Higgs pair searches as compared to SM measurements.

Searches for the production of Higgs pairs
have already been performed at Run I with 8 TeV
by ATLAS in the $b\bar{b}b\bar{b}$~\cite{Aad:2015uka}
and $b\bar{b}\gamma\gamma$~\cite{Aad:2014yja} final states,
and by
CMS in the same $b\bar{b}b\bar{b}$~\cite{Khachatryan:2015yea}
and $b\bar{b}\gamma\gamma$~\cite{Chatrchyan:2011wt} final
states.
%
%
In addition, ATLAS has presented~\cite{Aad:2015xja} a combination
of its di-Higgs searches in the $bb\tau\tau,$
$\gamma\gamma WW^*$, $\gamma\gamma bb$, and $bbbb$ final states.
%
Many other exotic searches involves Higgs pairs in the final
state, such as the recent
searches for heavy Higgs bosons $H$~\cite{Khachatryan:2015tha}.

In the context of SM production,
the main advantage of the $b\bar{b}b\bar{b}$ final state is the
enhancement of the signal yield
from the large branching fraction of Higgs bosons into bottom-anti-bottom
pairs, ${\rm BR}\lp H\to b\bar{b}\rp\simeq 0.57$~\cite{Dittmaier:2012vm}.
%
However, a measurement in this channel
needs to deal with an overwhelming QCD multi-jet background.
%
Recent studies of Higgs pair production in this
final state~\cite{Wardrope:2014kya,deLima:2014dta}
estimate that, for an integrated
luminosity of
$\mathcal{L}=3$ ab$^{-1}$,
a signal significance of around $S/\sqrt{B}\sim 2.0$ could be obtained.
%
While irreducible backgrounds, such as $4b$ and
$t\bar{t}$, are included in these analysis, the contribution
from reducible processes, such as $bbjj$ and
$jjjj$, which can contribute to the signal yield when 
light jets are mis-identified as $b$-jets,
are not accounted for.
%
As will show, due to both
selection effects and $b$-quark radiation in the
parton shower, the
contribution of the $2b2j$ process is actually comparable to
the irreducible $4b$ component.

>>>>

Analogously to single Higgs production~\cite{Dittmaier:2012vm}, 
in the SM the dominant mechanism for the production of a pair of
Higgs bosons at the LHC is 
gluon fusion (see Refs.~\cite{baglio,Frederix:2014hta} and
references therein).
%
For a center-of-mass energy $\sqrt{s} = 14\,$TeV, the
next-to-next-to-leading order (NNLO)
total cross section is approximately $40\,$fb~\cite{deFlorian:2013jea},
which is increased by a further few percent once
next-to-next-to-leading logarithmic
(NNLL) corrections
are accounted for~\cite{deFlorian:2015moa}.
%
Feasibility studies in the case of a SM-like Higgs boson
in the gluon-fusion channel
at the LHC have been performed for different final states, including
$b\bar b\gamma\gamma$~\cite{Baur:2003gp,Barger:2013jfa},
$b\bar{b}\tau^+\tau^-$~\cite{Baur:2003gpa,Barr:2013tda,Dolan:2012rv,Dolan:2013rja},
$b\bar{b}W^+W^-$~\cite{Dolan:2012rv,Papaefstathiou:2012qe} and
$b\bar{b}b\bar{b}$~\cite{Baur:2003gpa,Dolan:2012rv,Wardrope:2014kya,deLima:2014dta,Barger:2013jfa}.
%
While these studies differ in their quantitative conclusions,
a consistent picture emerges 
that the ultimate precision in the determination of the Higgs trilinear
coupling $\lambda$ requires the full integrated luminosity
of the HL-LHC, $\mathcal{L}=3$ ab$^{-1}$,
and will rely on the combination of all relevant final states.

@@

While the SM production rates for Higgs
pairs are small, these can be substantially
enhanced in a variety of BSM scenarios.
%
Feasibility studies of Higgs pair production in New Physics
models have been performed in a number of different frameworks,
including Effective Field
Theories (EFTs) with higher-dimensional
operators and anomalous 
Higgs couplings~\cite{Nishiwaki:2013cma,Dall'Osso:2015aia,Azatov:2015oxa,Liu:2014rba,Goertz:2014qta,He:2015spf,Grober:2015cwa}, resonant production
in models such as extra dimensions~\cite{Gouzevitch:2013qca,Cooper:2013kia,No:2013wsa,Wen-Juan:2015gqg}, and Supersymmetry and
Two Higgs Doublet models (2HDMs)~\cite{Belyaev:1999kk,Han:2013sga,Hespel:2014sla,Wu:2015nba,Cao:2014kya,Ellwanger:2013ova}.
%
Since BSM dynamics modify
the kinematic distributions of the Higgs decay products, for
instance boosting the di-Higgs system,
different analysis strategies  might be required for BSM
Higgs pair searches as compared to SM measurements.

Searches for the production of Higgs pairs
have already been performed with 8 TeV Run I data
by ATLAS in the $b\bar{b}b\bar{b}$~\cite{Aad:2015uka}
and $b\bar{b}\gamma\gamma$~\cite{Aad:2014yja} final states,
and by
CMS in the same $b\bar{b}b\bar{b}$~\cite{Khachatryan:2015yea}
and $b\bar{b}\gamma\gamma$~\cite{Chatrchyan:2011wt} final
states.
%
%
In addition, ATLAS has presented~\cite{Aad:2015xja} a combination
of its di-Higgs searches in the $bb\tau\tau,$
$\gamma\gamma WW^*$, $\gamma\gamma bb$, and $bbbb$ final states.
%
Many other exotic searches involves Higgs pairs in the final
state, such as the recent
searches for heavy Higgs bosons $H$~\cite{Khachatryan:2015tha}.

In the context of SM production,
the main advantage of the $b\bar{b}b\bar{b}$ final state is the
enhancement of the signal yield
from the large branching fraction of Higgs bosons into bottom-anti-bottom
pairs, ${\rm BR}\lp H\to b\bar{b}\rp\simeq 0.57$~\cite{Dittmaier:2012vm}.
%
However, a measurement in this channel
needs to deal with an overwhelming QCD multi-jet background.
%
Recent studies of Higgs pair production in this
final state~\cite{Wardrope:2014kya,deLima:2014dta}
estimate that, for an integrated
luminosity of
$\mathcal{L}=3$ ab$^{-1}$,
a signal significance of around $S/\sqrt{B}\sim 2.0$ could be obtained.
%
While irreducible backgrounds, such as $4b$ and
$t\bar{t}$, are included in these analysis, the contribution
from reducible processes, such as $bbjj$ and
$jjjj$, which can contribute to the signal yield when 
light jets are mis-identified as $b$-jets,
are not accounted for.
%
As will show, due to both
selection effects and $b$-quark radiation in the
parton shower, the
contribution of the $2b2j$ process is comparable to
the irreducible $4b$ component.
